Federated Learning's Blessing:\\ FedAvg has Linear Speedup

### 2020-09-25 source-iclr/app_sgd_scvx_smooth.tex
Update on Overleaf.

One may ask whether the dependence on $E$ in the term $\frac{\kappa E^{2}G^{2}/\mu}{KT}$
can be removed, or equivalently whether $\sum_{k}p_{k}\|\mathbf{w}_{t}^{k}-\overline{\mathbf{w}}_{t}\|^{2}=\mathcal{O}(1/T^{2})$
can be independent of $E$. We provide a simple counterexample that
shows that this is not possible in general. 
\begin{proposition}
	There exists a dataset such that if $E=\mathcal{O}(T^{\beta})$ for
	any $\beta>0$ then $\sum_{k}p_{k}\|\mathbf{w}_{t}^{k}-\overline{\mathbf{w}}_{t}\|^{2}=\Omega(\frac{1}{T^{2-2\beta}})$
	.
\end{proposition}
\begin{proof}
	Suppose that we have an even number of devices and each $F_{k}(\mathbf{w})=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}(\mathbf{x}_{k}^{j}-\mathbf{w})^{2}$
	contains data points $\mathbf{x}_{k}^{j}=\mathbf{w}^{\ast,k}$, with
	$n_{k}\equiv n$. Moreover, the $\mathbf{w}{}^{\ast,k}$'s come in
	pairs around the origin. As a result, the global objective $F$ is
	minimized at $\mathbf{w}^{\ast}=0$. Moreover, if we start from $\overline{\mathbf{w}}_{0}=0$,
	then by design of the dataset the updates in local steps exactly cancel
	each other at each iteration, resulting in $\overline{\mathbf{w}}_{t}=0$
	for all $t$. On the other hand, if $E=T^{\beta}$, then starting
	from any $t=\mathcal{O}(T)$ with constant step size $\mathcal{O}(\frac{1}{T})$,
	after $E$ iterations of local steps, the local parameters are updated
	towards $\mathbf{w}^{\ast,k}$ with $\|\mathbf{w}_{t+E}^{k}\|^{2}=\Omega((T^{\beta}\cdot\frac{1}{T})^{2})=\Omega(\frac{1}{T^{2-2\beta}})$.
	This implies that 
	\begin{align*}
	\sum_{k}p_{k}\|\mathbf{w}_{t+E}^{k}-\overline{\mathbf{w}}_{t+E}\|^{2} & =\sum_{k}p_{k}\|\mathbf{w}_{t+E}^{k}\|^{2}\\
	& =\Omega(\frac{1}{T^{2-2\beta}})
	\end{align*}
	which is at a slower rate than $\frac{1}{T^{2}}$ for any $\beta>0$.
	Thus the sampling variance $\mathbb{E}\|\overline{\mathbf{w}}_{t+1}-\overline{\mathbf{v}}_{t+1}\|^{2}=\Omega(\sum_{k}p_{k}\mathbb{E}\|\mathbf{w}_{t+1}^{k}-\overline{\mathbf{w}}_{t+1}\|^{2})$
	decays at a slower rate than $\frac{1}{T^{2}}$, resulting in a convergence
	rate slower than $\mathcal{O}(\frac{1}{T})$ with partial participation. 
\end{proof}

>>>>

One may ask whether the dependence on $E$ in the term $\frac{\kappa E^{2}G^{2}/\mu}{KT}$
can be removed, or equivalently whether $\sum_{k}p_{k}\|\mathbf{w}_{t}^{k}-\overline{\mathbf{w}}_{t}\|^{2}=\mathcal{O}(1/T^{2})$
can be independent of $E$. We provide a simple counterexample that
shows that this is not possible in general. 
\begin{proposition} \label{prop:tight}
	There exists a dataset such that if $E=\mathcal{O}(T^{\beta})$ for
	any $\beta>0$ then $\sum_{k}p_{k}\|\mathbf{w}_{t}^{k}-\overline{\mathbf{w}}_{t}\|^{2}=\Omega(\frac{1}{T^{2-2\beta}})$
	.
\end{proposition}
\begin{proof}
	Suppose that we have an even number of devices and each $F_{k}(\mathbf{w})=\frac{1}{n_{k}}\sum_{j=1}^{n_{k}}(\mathbf{x}_{k}^{j}-\mathbf{w})^{2}$
	contains data points $\mathbf{x}_{k}^{j}=\mathbf{w}^{\ast,k}$, with
	$n_{k}\equiv n$. Moreover, the $\mathbf{w}{}^{\ast,k}$'s come in
	pairs around the origin. As a result, the global objective $F$ is
	minimized at $\mathbf{w}^{\ast}=0$. Moreover, if we start from $\overline{\mathbf{w}}_{0}=0$,
	then by design of the dataset the updates in local steps exactly cancel
	each other at each iteration, resulting in $\overline{\mathbf{w}}_{t}=0$
	for all $t$. On the other hand, if $E=T^{\beta}$, then starting
	from any $t=\mathcal{O}(T)$ with constant step size $\mathcal{O}(\frac{1}{T})$,
	after $E$ iterations of local steps, the local parameters are updated
	towards $\mathbf{w}^{\ast,k}$ with $\|\mathbf{w}_{t+E}^{k}\|^{2}=\Omega((T^{\beta}\cdot\frac{1}{T})^{2})=\Omega(\frac{1}{T^{2-2\beta}})$.
	This implies that 
	\begin{align*}
	\sum_{k}p_{k}\|\mathbf{w}_{t+E}^{k}-\overline{\mathbf{w}}_{t+E}\|^{2} & =\sum_{k}p_{k}\|\mathbf{w}_{t+E}^{k}\|^{2}\\
	& =\Omega(\frac{1}{T^{2-2\beta}})
	\end{align*}
	which is at a slower rate than $\frac{1}{T^{2}}$ for any $\beta>0$.
	Thus the sampling variance $\mathbb{E}\|\overline{\mathbf{w}}_{t+1}-\overline{\mathbf{v}}_{t+1}\|^{2}=\Omega(\sum_{k}p_{k}\mathbb{E}\|\mathbf{w}_{t+1}^{k}-\overline{\mathbf{w}}_{t+1}\|^{2})$
	decays at a slower rate than $\frac{1}{T^{2}}$, resulting in a convergence
	rate slower than $\mathcal{O}(\frac{1}{T})$ with partial participation. 
\end{proof}

### 2020-09-25 source-iclr/main.tex
iclr from nips version 1.
update exp
add scaffold and analysis
proof th1 minor
Update main.tex
Update on Overleaf.

% ready for submission
\usepackage{neurips_2020}

@@

% to avoid loading the natbib package, add option nonatbib:
     % \usepackage[nonatbib]{neurips_2020}
\usepackage{slashbox,multirow}
% \usepackage{tikz}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amsthm}     % for theorems
\usepackage{algorithm,algorithmic}
\usepackage{mathtools}
\usepackage{amsmath}
\allowdisplaybreaks
\usepackage{amssymb}
\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
\usepackage{balance}
\usepackage{comment}
\newcommand{\eq}[1]{{Eq~(#1)}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newtheorem{theorem}{Theorem}
\newtheorem{thm}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{remark}{Remark}
\newcommand{\lkxcom}[1]{{\color{red}{#1}}}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\def\lowcomma{_{\textstyle{,}}}
\def\lowperiod{_{\textstyle.}}

\newcommand{\ov}[1]{{\overline{\mathbf{#1}}}}
\input{bmacros}
\newtheorem*{assumption*}{Assumption}

@@

\begin{abstract}
Federated learning (FL) learns a model jointly from a set of participating devices without sharing each other's privately held data. The characteristics of non-\textit{iid} data across the network, low device participation, and the mandate that data remain private bring challenges in understanding the convergence of FL algorithms, particularly in regards to how convergence scales with the number of participating devices. In this paper, we focus on Federated Averaging (FedAvg)--the most widely used and effective FL algorithm in use today--and provide a comprehensive study of its convergence rate. Although FedAvg has recently been studied by an emerging line of literature, it remains open as to how FedAvg's convergence scales with the number of participating devices in the FL setting--a crucial question whose answer would shed light on the performance of FedAvg in large FL systems. We fill this gap by establishing convergence guarantees for FedAvg under three classes of problems: strongly convex smooth, convex smooth, and overparameterized strongly convex smooth problems. We show that FedAvg enjoys linear speedup in each case, although with different convergence rates. For each class, we also characterize the corresponding convergence rates for the Nesterov accelerated FedAvg algorithm in the FL setting: to the best of our knowledge, these are the first linear speedup guarantees for FedAvg when Nesterov acceleration is used. To accelerate FedAvg, we also design a new momentum-based FL algorithm that further improves the convergence rate in overparameterized linear regression problems. Empirical studies of the algorithms in various settings have supported our theoretical results.
\end{abstract}

% (and its related variant local SGD also analyzed in the simpler homogeneous distributed optimization setting)
\input{sec_intro}
%\input{sec_relatedwork}
\input{sec_problem}
\input{sec_SGD}
\input{sec_Nesterov}
\input{sec_overparameterized}
\input{sec_exp}

@@

% \newpage
\input{sec_impact}
\bibliographystyle{unsrt}
\bibliography{ref.bib}

>>>>

% ready for submission
\usepackage{iclr2021_conference,times}

@@

% to avoid loading the natbib package, add option nonatbib:
     % \usepackage[nonatbib]{neurips_2020}
\usepackage{slashbox,multirow}
% \usepackage{tikz}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amsthm}     % for theorems
\usepackage{algorithm,algorithmic}
\usepackage{mathtools}
\usepackage{amsmath}
\allowdisplaybreaks
\usepackage{amssymb}
\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
\usepackage{balance}
\usepackage{comment}
\newcommand{\eq}[1]{{Eq~(#1)}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newtheorem{theorem}{Theorem}
\newtheorem{thm}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{remark}{Remark}
\newtheorem{proposition}{Proposition}
\newcommand{\lkxcom}[1]{{\color{red}{#1}}}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\def\lowcomma{_{\textstyle{,}}}
\def\lowperiod{_{\textstyle.}}

\newcommand{\ov}[1]{{\overline{\mathbf{#1}}}}
\input{bmacros}

\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}

\newtheorem*{assumption*}{Assumption}
\newtheorem*{claim*}{Claim}

@@

\begin{abstract}
\small Federated learning (FL) learns a model jointly from a set of participating devices without sharing each other's privately held data. The characteristics of non-\textit{i.i.d.} data across the network, low device participation, high communication costs, and the mandate that data remain private bring challenges in understanding the convergence of FL algorithms, particularly in regards to how convergence scales with the number of participating devices. In this paper, we focus on Federated Averaging (FedAvg)--arguably the most popular and effective FL algorithm class in use today--and provide a unified and comprehensive study of its convergence rate. Although FedAvg has recently been studied by an emerging line of literature, it remains open as to how FedAvg's convergence scales with the number of participating devices in the fully heterogeneous FL setting--a crucial question whose answer would shed light on the performance of FedAvg in large FL systems. We fill this gap by providing a unified analysis that establishes convergence guarantees for FedAvg under three classes of problems: strongly convex smooth, convex smooth, and overparameterized strongly convex smooth problems. We show that FedAvg enjoys linear speedup in each case, although with different convergence rates and communciation efficiencies. While there have been linear speedup results from distributed optimization that assumes full participation, ours are the first to establish linear speedup for FedAvg under both statistical and system heterogeneity.  For strongly convex and convex problems, we also characterize the corresponding convergence rates for the Nesterov accelerated FedAvg algorithm, which are the first linear speedup guarantees for momentum variants of FedAvg in the convex setting. To provably accelerate FedAvg, we design a new momentum-based FL algorithm that further improves the convergence rate in overparameterized linear regression problems. Empirical studies of the algorithms in various settings have supported our theoretical results.
\end{abstract}

% (and its related variant local SGD also analyzed in the simpler homogeneous distributed optimization setting)
\input{sec_intro}
%\input{sec_relatedwork}
\input{sec_problem}
\input{sec_SGD}
\input{sec_Nesterov}
%\input{sec_overparameterized}
\input{sec_exp}

@@

% \newpage
% \input{sec_impact}
\newpage
\bibliographystyle{iclr2021_conference}
\bibliography{ref.bib}

### 2020-09-25 source-iclr/sec_Nesterov.tex
Update on Overleaf.

A natural extension of the FedAvg algorithm is to use momentum-based
local updates instead of local SGD updates. To our knowledge, the
only convergence analyses of FedAvg with momentum-based stochastic
updates focus on the non-convex smooth case \cite{huo2020faster,yu2019linear,li2018federated}.
In this section, we complete the picture with $\mathcal{O}(1/NT)$
and $\mathcal{O}(1/\sqrt{NT})$ convergence results for Nesterov-accelerated
FedAvg for convex objectives that match the rates from the previous section.
As we know from stochastic optimization, Nesterov and other momentum
updates may fail to accelerate over SGD \cite{liu2018accelerating,kidambi2018insufficiency,liu2018toward,yuan2016influence}.
Therefore in Section~\ref{sec:overparameterized} we will specialize to overparameterized
problems where we demonstrate that a particular FedAvg variant with
momentum updates is able to accelerate over the original FedAvg algorithm. 
Detailed proofs of convergence results in this section are deferred to Appendix Section~\ref{sec:app:Nesterovfedavg}.

\subsection{Strongly Convex and Smooth Objectives}
\begin{comment}
We first show that the Nesterov accelerated FedAvg has $\mathcal{O}(1/NT)$
convergence rate for $\mu$-strongly convex and $L$-smooth objectives.
\end{comment}
The Nesterov Accelerated FedAvg algorithm follows the updates:
\begin{align*}
\mathbf{v}_{t+1}^{k} & =\mathbf{w}_{t}^{k}-\alpha_{t}\mathbf{g}_{t,k}, \hspace{1em}
\mathbf{w}_{t+1}^{k} =\begin{cases}
\mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k}) & \text{if }t+1\notin\mathcal{I}_{E},\\
\sum_{k \in \cS_{t+1}}q_k\left[\mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k})\right] & \text{if }t+1\in\mathcal{I}_{E},
\end{cases}
\end{align*}
where $\mathbf{g}_{t,k}:=\nabla F_{k}(\mathbf{w}_{t}^{k},\xi_{t}^{k})$ is
the stochastic gradient sampled on the $k$-th device at time $t$, and $q_k$ again depends on participation and sampling schemes.  
\begin{theorem}
	\label{thm:nesterov_scvx}Let $\overline{\mathbf{v}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{v}_{T}^{k}$ in Nesterov accelerated FedAvg,
	and set learning rates $\alpha_{t}=\frac{6}{\mu}\frac{1}{t+\gamma}$,  $\beta_{t-1}=\frac{3}{14(t+\gamma)(1-\frac{6}{t+\gamma})\max\{\mu,1\}}$. Then under Assumptions~\ref{ass:lsmooth},\ref{ass:stroncvx},\ref{ass:boundedvariance},\ref{ass:subgrad2} with full device participation, 
	\begin{align*}
	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right),
	\end{align*}
	and with partial device participation with $K$ sampled devices at
	each communication round, 
	\begin{align*}
	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right).
	\end{align*}
\end{theorem}
%
To our knowledge, this is the first convergence result for Nesterov
accelerated FedAvg in the strongly convex and smooth setting. The
same discussion about linear speedup of FedAvg applies to the Nesterov
accelerated variant. In particular, to achieve $\mathcal{O}(1/NT)$
linear speedup, $T$ iterations of the algorithm require only $\mathcal{O}(\sqrt{NT})$
communication rounds with full participation.

@@

We now show that the optimality gap of Nesterov Accelerated FedAvg has $\mathcal{O}(1/\sqrt{NT})$ rate. This result complements the strongly convex case in the previous
part, as well as the non-convex smooth setting in \cite{huo2020faster,yu2019linear,li2018federated},
where a similar $\mathcal{O}(1/\sqrt{NT})$ rate is given in terms
of averaged gradient norm. 
\begin{theorem}
	\label{thm:Nesterov_cvx}Set learning rates $\alpha_{t}=\beta_{t}=\mathcal{O}(\sqrt{\frac{N}{T}})$. Then under Assumptions~\ref{ass:lsmooth},\ref{ass:boundedvariance},\ref{ass:subgrad2} Nesterov accelerated FedAvg with
	full device participation has rate
	\begin{align*}
	\min_{t\leq T}F(\overline{\mathbf{v}}_{t})-F^{\ast} & =\mathcal{O}\left(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T}\right),
	\end{align*}
	and with partial device participation with $K$ sampled devices at
	each communication round, 
	\begin{align*}
	\min_{t\leq T}F(\overline{\mathbf{v}}_{t})-F^{\ast} & =\mathcal{O}\left(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T}\right).
	\end{align*}
\end{theorem}
%
	It is possible to extend the results in this section to accelerated
    FedAvg algorithms with other momentum-based updates. However, in the
    stochastic optimization setting, none of these methods can achieve a better rate than the original FedAvg with SGD updates for general problems~\cite{kidambi2018insufficiency}.
    For this reason, we will instead turn to the overparameterized setting
    \cite{ma2017power,liu2018accelerating,canziani2016analysis} in the
    next section where we show that FedAvg enjoys geometric convergence
    and it is possible to improve its convergence rate with momentum-based updates.

>>>>

A natural extension of the FedAvg algorithm is to use momentum-based
local updates instead of local SGD updates. To our knowledge, the
only convergence analyses of FedAvg with momentum-based stochastic
updates focus on the non-convex smooth case \cite{huo2020faster,yu2019linear,li2018federated}.
In this section, we complete the picture by providing the first $\mathcal{O}(1/KT)$
and $\mathcal{O}(1/\sqrt{KT})$ convergence results for Nesterov-accelerated
FedAvg for convex objectives that match the rates from the previous section. Detailed proofs of convergence results in this section are deferred to Appendix Section~\ref{sec:app:Nesterovfedavg}.

\subsection{Strongly Convex and Smooth Objectives}
\begin{comment}
We first show that the Nesterov accelerated FedAvg has $\mathcal{O}(1/NT)$
convergence rate for $\mu$-strongly convex and $L$-smooth objectives.
\end{comment}
The Nesterov Accelerated FedAvg algorithm follows the updates:
\begin{align*}
\mathbf{v}_{t+1}^{k} & =\mathbf{w}_{t}^{k}-\alpha_{t}\mathbf{g}_{t,k}, \hspace{1em}
\mathbf{w}_{t+1}^{k} =\begin{cases}
\mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k}) & \text{if }t+1\notin\mathcal{I}_{E},\\
\sum_{k \in \cS_{t+1}}q_k\left[\mathbf{v}_{t+1}^{k}+\beta_{t}(\mathbf{v}_{t+1}^{k}-\mathbf{v}_{t}^{k})\right] & \text{if }t+1\in\mathcal{I}_{E},
\end{cases}
\end{align*}
where $\mathbf{g}_{t,k}:=\nabla F_{k}(\mathbf{w}_{t}^{k},\xi_{t}^{k})$ is
the stochastic gradient sampled on the $k$-th device at time $t$, and $q_k$ again depends on participation and sampling schemes.  
\begin{theorem}
	\label{thm:nesterov_scvx}Let $\overline{\mathbf{v}}_{T}=\sum_{k=1}^{N}p_{k}\mathbf{v}_{T}^{k}$ in Nesterov accelerated FedAvg,
	and set learning rates $\alpha_{t}=\frac{6}{\mu}\frac{1}{t+\gamma}$,  $\beta_{t-1}=\frac{3}{14(t+\gamma)(1-\frac{6}{t+\gamma})\max\{\mu,1\}}$. Then under Assumptions~\ref{ass:lsmooth},\ref{ass:stroncvx},\ref{ass:boundedvariance},\ref{ass:subgrad2} with full device participation, 
	\begin{align*}
	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right),
	\end{align*}
	and with partial device participation with $K$ sampled devices at
	each communication round, 
	\begin{align*}
	\mathbb{E}F(\overline{\mathbf{v}}_{T})-F^{\ast}=\mathcal{O}\left(\frac{\kappa\nu_{\max}^{2}\sigma^{2}/\mu}{NT}+\frac{\kappa E^{2}G^{2}/\mu}{KT}+\frac{\kappa^{2}E^{2}G^{2}/\mu}{T^{2}}\right).
	\end{align*}
\end{theorem}
The key step in the proof of this result is a one step contraction that involves three time steps, due to the update format of Nesterov SGD, but shares similar elements as that for FedAvg:
\begin{align*}
\mathbb{E}\|\ov{v}_{t+1}-\vw^{\ast}\|^{2} & \leq\mathbb{E}(1-\mu\alpha_{t})(1+\beta_{t-1})^{2}\|\ov{v}_{t}-\vw^{\ast}\|^{2}+20E^{2}L\alpha_{t}^{3}G^{2}+(1-\alpha_{t}\mu)\beta_{t-1}^{2}\|(\ov{v}_{t-1}-\vw^{\ast})\|^{2}\\
& +\alpha_{t}^{2}\frac{1}{N}\nu_{\max}\sigma^{2}+2\beta_{t-1}(1+\beta_{t-1})(1-\alpha_{t}\mu)\|\ov{v}_{t}-\vw^{\ast}\|\cdot\|\ov{v}_{t-1}-\vw^{\ast}\|
\end{align*}
%
To our knowledge, this is the first convergence result for Nesterov
accelerated FedAvg in the strongly convex and smooth setting. The
same discussion about linear speedup of FedAvg applies to the Nesterov
accelerated variant. In particular, to achieve $\mathcal{O}(1/NT)$
linear speedup, $T$ iterations of the algorithm require only $\mathcal{O}(\sqrt{NT})$
communication rounds with full participation.

@@

We now show that the optimality gap of Nesterov Accelerated FedAvg has $\mathcal{O}(1/\sqrt{KT})$ rate. This result complements the strongly convex case in the previous
part, as well as the non-convex smooth setting in \cite{huo2020faster,yu2019linear,li2018federated},
where a similar $\mathcal{O}(1/\sqrt{KT})$ rate is given in terms
of averaged gradient norm. 
\begin{theorem}
	\label{thm:Nesterov_cvx}Set learning rates $\alpha_{t}=\beta_{t}=\mathcal{O}(\sqrt{\frac{N}{T}})$. Then under Assumptions~\ref{ass:lsmooth},\ref{ass:boundedvariance},\ref{ass:subgrad2} Nesterov accelerated FedAvg with
	full device participation has rate
	\begin{align*}
	\min_{t\leq T}F(\overline{\mathbf{v}}_{t})-F^{\ast} & =\mathcal{O}\left(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{NT}}+\frac{NE^{2}LG^{2}}{T}\right),
	\end{align*}
	and with partial device participation with $K$ sampled devices at
	each communication round, 
	\begin{align*}
	\min_{t\leq T}F(\overline{\mathbf{v}}_{t})-F^{\ast} & =\mathcal{O}\left(\frac{\nu_{\max}^{2}\sigma^{2}}{\sqrt{KT}}+\frac{E^{2}G^{2}}{\sqrt{KT}}+\frac{KE^{2}LG^{2}}{T}\right).
	\end{align*}
\end{theorem}
%
	It is possible to extend the results in this section to accelerated
    FedAvg algorithms with other momentum-based updates. However, as know from stochastic optimization, Nesterov and other momentum
updates may fail to accelerate over SGD (\cite{liu2018accelerating,kidambi2018insufficiency,liu2018toward,yuan2016influence}).
    For this reason, we will instead turn to the overparameterized setting
    \cite{ma2017power,liu2018accelerating,canziani2016analysis} in Section~\ref{sec:app:overparameterized} of the appendix where we show that FedAvg enjoys geometric convergence
    and it is possible to improve its convergence rate with a new momentum-based FedAvg variant, which we term ``FedMaSS''.
